---
title: "Measuring Child Poverty Around The World"
format: html
execute:
  echo: false
  warning: false
  message: false
---



Children suffer poverty differently from adults. Their needs, to survive and thrive are different. Children are not supposed to earn a living for themselves. Thus, when measuring child poverty, it is important to use a direct multidimensional measurement. Due to data limitations, only the following dimensions were estimated: <span style="color:blue;">education, health, housing, nutrition, sanitation, and water </span>



## Average Percentage of Child Poverty by Gender

The following chart depicts the Average percentage of male and female children suffering from at least two deprivation between the periods 2011-2018. The <span style="color:pink;"> total percentage of female children suffering from at least 2 deprivation is 36.92% </span> and with a very minimal difference, <span style="color:blue;"> the total percentage of male children suffering from child deprivation is 37.28% </span>. Despite the minor difference, both genders face child deprivation at similar rates. This also indicates that the social factors that impact child poverty are fairly distributed among both genders.


```{r}
install.packages("tidyverse")
install.packages("ggplot2")
install.packages("maps")

# Load necessary libraries
library(tidyverse)
library(ggplot2)

# Read the dataset
data <- read.csv("/cloud/project/unicef_indicator_1.csv")
metadata <- read.csv("/cloud/project/unicef_metadata.csv")

#print(colnames(metadata))

# Filter the data to include only 'Male' and 'Female'
gender_data <- data %>%
  filter(sex %in% c("Male", "Female"))

# Group by gender and summarize to find the average percentage of child poverty
gender_averages <- gender_data %>%
  group_by(sex) %>%
  summarize(average_percentage = mean(obs_value, na.rm = TRUE))

# Create the bar plot for average percentages
ggplot(gender_averages, aes(x = sex, y = average_percentage, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = sprintf("%.2f%%", average_percentage)), vjust = -0.5, color = "black", size = 3.5) +
  labs(x = "Gender",
       y = "Average Percentage") +
  theme_minimal()
```

## World Map of Child Poverty by Country and Time Period

The rate of children suffering from 2 deprivation in different countries from 2011 to 2018 is shown in the following world map. In comparison,<span style="color:red;">Africa is the most affected country and the rate has been higher throughout the years</span>. Policymakers and international organizations should continue to prioritize interventions and aid efforts in Africa to address these issues.

```{r}
# Calculate average percentage of child poverty per country and time period
average_poverty <- data %>%
  group_by(country, time_period) %>%
  summarize(average_poverty = mean(obs_value, na.rm = TRUE), .groups = 'drop')

# Get world map data
world_map <- map_data("world")

# Merge your data with the world map data
world_data <- world_map %>%
  left_join(average_poverty, by = c("region" = "country"))

# Plot the data
ggplot(data = world_data, aes(x = long, y = lat, group = group)) +
  geom_polygon(aes(fill = average_poverty)) +
  scale_fill_gradient(low = "pink", high = "red", name = "Average % of Child Poverty") +
  coord_fixed(1.3) +
  labs(fill = "Average % of Child Poverty") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Scatterplot of Life Expectancy vs. Child Poverty

Through this scatterplot, we can understand how the child poverty rate affects the life expectancy of a human being. It is evident that the lower the child poverty rate the higher life expectancy a person has and vice versa,<span style="color:orange;">it has an inversely proportional relationship </span> . For example, if the average percentage of child poverty is 2.74 then their life expectancy is 80 years and if the average percentage of child poverty is 81 then their life expectancy is 51 years.


```{r}

filtered_data <- read.csv("/cloud/project/FilteredData.csv")

#print(colnames(filtered_data))

# Correct column names by replacing all dots and excess punctuation
names(filtered_data) <- names(filtered_data) %>%
  str_replace_all("\\.", " ") %>%
  str_replace_all(" {2,}", " ") %>%
  str_trim()

# Make sure column names are corrected by printing them
#print(names(filtered_data))

# Filter by sex 
filter_data <- filtered_data %>%
  filter(sex %in% c("Male", "Female"))

#print(names(filter_data))

ggplot(filter_data, aes(x = `Life expectancy at birth total years`, y = obs_value, color = sex)) +
  geom_point() +
  labs(x = "Life Expectancy at Birth (years)",
    y = "Average Percentage of Child Poverty",
    caption = "Data includes both Male and Female") +
  theme_minimal() +
  scale_color_manual(values = c("Male" = "yellow", "Female" = "orange"))
```

## Time Series of GNI for Top 10 Countries

The time series chart displays the Gross National Income (GNI) of the top 10 Countries between <span style="color:red;">1960 and 2020 </span>.GNI is a metric that measures a country's economic output. According to the chart, China has the highest GNI growth rate, while Brazil has the lowest growth rate among the top ten countries.


```{r}


# Scatter plot for GNI on top 10 countries 

# Correct column names by replacing all dots and excess punctuation
names(metadata) <- names(metadata) %>%
  str_replace_all("\\.", " ") %>%
  str_replace_all(" {2,}", " ") %>%
  str_trim()

#print(colnames(metadata))

# List of top 10 countries
top_countries <- c("United States", "China", "Japan", "Germany", "United Kingdom", 
                   "France", "Brazil", "Italy", "Russia", "Canada")

# Prepare and plot the data for top 10 countries
metadata %>%
  filter(!is.na(`GNI current US`) & country %in% top_countries) %>%  # Filter for top 10 countries and non-missing GNI values
  mutate(year = as.Date(paste(year, "01", "01", sep = "-"))) %>%  # Create a proper date column
  ggplot(aes(x = year, y = `GNI current US`, group = country, color = country)) +  # Plotting GNI over time for each country
  geom_line() +  # Use line type for time series
  scale_y_continuous(labels = scales::unit_format(unit = "B", scale = 1e-9),  # Format y-axis labels with "b" for billion
                     breaks = c(0, 1000e9, 2000e9, 3000e9, 4000e9),  # Define breaks for y-axis in billions
                     limits = c(0, 4000e9),  # Set limits for y-axis
                     name = "GNI (current US$)") +  # Label for y-axis
  labs(x = "Year",
       y = "GNI (current US$)",
       color = "Country") +
  theme_minimal() +
  theme(legend.position = "bottom")  # Adjust legend position
```

## Conclusion

Measuring child poverty requires a multifaceted approach that takes calculates factors such as education, health, and housing access. Despite minor gender differences, both male and female children suffer from deprivation at the same rate, indicating that social factors are evenly distributed. Africa emerges as the most affected region, necessitating priority interventions to address ongoing challenges. The inverse relationship between child poverty and life expectancy emphasizes the critical importance of addressing socioeconomic disparities to ensure long-term well-being. Understanding economic trends, such as GNI growth rates, helps to inform global efforts to combat child poverty and promote sustainable development.
